Just Add IFC! A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-03-02 paper/real.tex

\subsection{C}
\label{sec:real:c}
%
C programs are able to execute arbitrary (machine) code, access
arbitrary memory, and perform arbitrary syscalls.
%
Thus, the confinement of C programs must be imposed by the underlying OS
and hardware.
%
This isolation can be achieved using hardware protection
mechanisms~\cite{Belay:2012:DSU:2387880.2387913} (similar to a simpler
version of Wedge~\cite{Bittau:2008:WSA:1387589.1387611} with an
information flow control policy).
%
Using pages tables, a (trusted) IFC runtime could ensure that each task,
implemented as a lightweight process, can only access the memory it
allocates--tasks do not have access to any shared memory.
%
In addition, ring protection could be used to intercept syscalls performed by
a task and only permit those corresponding to our IFC language (e.g.,
|getLabel|, |send|).
%
Although other sandboxing mechanisms can be used in place of Dune, we
believe Dune's hardware protection mechanism would allow us to provide a 
concrete implementation that is more efficient and simpler to reason about.

In this setting, the combined language of Section~\ref{sec:retrofit}
can be interpreted in the following way: calling from the target
language to the IFC language corresponds to invoking a syscall.
%
Creating a new task with the |sandbox| syscall is corresponds to
\emph{forking} a process.  Here, page tables limit what memory
a task can access: we can ensure there will be no memory (effectively
defining |klone(tS)
= tS0|, where |tS0| is the set of pages necessary to bootstrap a
lightweight process.)
%
Similarly, control over page tables and protection bits allows us to
define a |send| syscall that copies pages to our
(trusted) runtime queue; and, correspondingly, a |recv| that copies
the pages from the runtime queue to the (untrusted) receiver.
%
Since C is not memory safe, conditions on these syscalls are
meaningless.
%
\Red{
EZY: Suggest chopping

    Of course, a more efficient message-passing approach, such as that of
Singularity~\tocite{}, which simply passes a pointer to data the
receiver at the cost of revoking the sender's access to the pages is
also possible.
%
However, since we do not consider the isomorphism condition for this,
we do not discuss it further.}

\subsection{Haskell}
\label{sec:real:hs}
In contrast with the C embedding which relies on hardware protection
mechanisms, a Haskell implementation can leverage Haskell's strong data
abstraction and static type system, monadic approach to effects, and
lightweight concurrency to implement the embedding in a more lightweight
manner.  We briefly describe one such system, LIO,~\cite{lio} which
implements IFC as a library.

LIO is implemented by defining a new monad, \verb|LIO|, which wraps Haskell's \verb|IO|
monad.
%
The purpose of this monad is twofold: it restricts the use of
arbitrary effects that would ordinarily be allowed by the \verb|IO| monad,
and it associates labels with threads (tasks).
%
Computations in the \verb|LIO|
monad can be thought to be operating within the IFC system.
%
One important aspect of using \verb|IO| as the base for this
implementation is that it allows use of Haskell's efficient implementation of
threads, channels, etc. (e.g., we use Haskell's \texttt{forkIO} to
fork a lightweight thread in the case of |fork|), in contrast to
defining them in a completely pure fashion (as suggested in Section~\ref{sec:monad}).

What is the interpretation of this system as per Section~\ref{sec:retrofit}?
%
Here, the \emph{pure subset} of Haskell is the target language
|targetLangHS|, while the monadic subset of Haskell in the \verb|LIO|
monad is the IFC language.
%
Crucially, the lack of unrestricted mutation in the pure fragment of
Haskell prevents direct communication between tasks, even when memory is
shared between them.\footnote{However, lazy evaluation can still produce
a covert channel.}

@@

We describe SWAPI,~\tocite{} an implementation of this system for JavaScript.
%
The direct approach for implementing the embedded language |specLangJS roundrobinf| is
by running multiple instances of the JS runtime, i.e., |targetLangJS|, in
separate OS-level threads.\footnote{
 The Firefox implementation of SWAPI also considers the case where
 tasks share the JavaScript runtime event-loop, but the heap and
 execution context of each task is disjoint. This is similar to the
 single-heap language of Section~\ref{sec:concrete}.  However, since the heap
 separation is provided by the browser the browser, we do not discuss
 this further. The interested reader is referred to~\tocite{} for more
 details.
}

From here, the IFC language functionality is implemented in the JS
runtime, much like browser layout engines implement the DOM, and
expose the new functions, e.g., |sandbox|, by attaching them to the JS
global object.
%
To avoid termination timing attacks, each task created with |sandbox| is
executed in a separate thread, running a separate, \emph{fresh} instance of our
modified JS runtime.
%
Formally, \verb|sandbox| is defined with |klone(tS) =
tS0|, where |tS0| is the global object corresponding to the standard
JS library (e.g., |tS0| contains \texttt{Object}, \texttt{Array},
etc.).

Since this implementation approach relies on multiple runtimes of the
language, sending arbitrary objects between tasks can result in
unexpected behavior, e.g., when the object contains references.
%
As discussed in Section~\ref{sec:concrete}, we need to restrict |send|
to expressions that can be marshalled as strings, i.e., structurally
clonable objects.
%
In our formalization, this amounts to restricting the IFC language rule
for |send| such that only strings can be shared:
%{
\newcommand{\str}{"string"}
%format tOf (e) = "\texttt{typeOf}("e")\texttt{ === \str}"
%format ttrue = "\texttt{true}"
\begin{mathpar}
\inferrule[JS-send]
{
|il canFlowTo il'|\\
|iS(id') = Q|\\
|iS' = iS [ id' mapsto (il', id,  ie') , Q ]|\\
|ie = IT te|\\
|conf tS (tOf te) -> conf tS ttrue|
}
{|
iconf iS (fullconf id il tS (iniEi (send id' il' ie)), ldots)
.->
iS'; sched S (fullconf id il tS (iniEi unit), ldots)
|}
\end{mathpar}
%}
We remark that this is similar to the existing
\texttt{postMessage} API used for iframe and Worker
communications~\cite{webworkers}.
%
Thus, the SWAPI implementation provides an IFC version of
\texttt{postMessage}, defined in terms of |send|, that additionally
takes a label argument.

>>>>

\subsection{C}
\label{sec:real:c}
%
C programs are able to execute arbitrary (machine) code, access
arbitrary memory, and perform arbitrary syscalls.
%
Thus, the confinement of C programs must be imposed by the underlying OS
and hardware.
%
This isolation can be achieved using Dune's hardware protection
mechanisms~\cite{Belay:2012:DSU:2387880.2387913}, similar to a simpler
version of Wedge~\cite{Belay:2012:DSU:2387880.2387913,
Bittau:2008:WSA:1387589.1387611} with an information flow control
policy.
%
Using pages tables, a (trusted) IFC runtime could ensure that each task,
implemented as a lightweight process, can only access the memory it
allocates--tasks do not have access to any shared memory.
%
In addition, ring protection could be used to intercept syscalls performed by
a task and only permit those corresponding to our IFC language (e.g.,
|getLabel|, |send|).
%
Although other sandboxing mechanisms can be used in place of Dune, we
believe Dune's hardware protection mechanism would allow us to provide a 
concrete implementation that is more efficient and simpler to reason about.

In this setting, the combined language of Section~\ref{sec:retrofit}
can be interpreted in the following way: calling from the target
language to the IFC language corresponds to invoking a syscall.
%
Creating a new task with the |sandbox| syscall is corresponds to
\emph{forking} a process.  Here, page tables limit what memory
a task can access: we can ensure there will be no memory (effectively
defining |klone(tS)
= tS0|, where |tS0| is the set of pages necessary to bootstrap a
lightweight process.)
%
Similarly, control over page tables and protection bits allows us to
define a |send| syscall that copies pages to our
(trusted) runtime queue; and, correspondingly, a |recv| that copies
the pages from the runtime queue to the (untrusted) receiver.
%
Since C is not memory safe, conditions on these syscalls are
meaningless.

\subsection{Haskell}
\label{sec:real:hs}
In contrast with the C embedding which relies on hardware protection
mechanisms, a Haskell implementation can leverage Haskell's strong data
abstraction and static type system, monadic approach to effects, and
lightweight concurrency to implement the embedding in a more lightweight
manner.  We briefly describe one such system, LIO~\cite{lio}, which
implements IFC as a library.

LIO is implemented by defining a new monad, \verb|LIO|, which wraps Haskell's \verb|IO|
monad.
%
The purpose of this monad is twofold: it restricts the use of
arbitrary effects that would ordinarily be allowed by the \verb|IO| monad,
and it associates labels with tasks.
%
Computations in the \verb|LIO|
monad can be thought to be operating within the IFC system.
%
One important aspect of using \verb|IO| as the base for this
implementation is that it allows use of Haskell's efficient
implementation of threads, channels, etc. (e.g., in the concurrent
version of LIO, we use Haskell's \texttt{forkIO} to fork a lightweight
thread in the case of |fork|~\cite{stefan:addressing-covert}), in
contrast to defining them in a completely pure fashion (as suggested
in Section~\ref{sec:monad}).

What is the interpretation of this system as per Section~\ref{sec:retrofit}?
%
Here, the \emph{pure subset} of Haskell is the target language, while
the monadic subset of Haskell in the \verb|LIO| monad is the IFC
language.
%
Crucially, the lack of unrestricted mutation in the pure fragment of
Haskell prevents direct communication between tasks, even when memory is
shared between them.\footnote{However, lazy evaluation can still produce
a covert channel.}

@@

We describe SWAPI~\tocite{}, an implementation of this system for JavaScript.
%
The direct approach for implementing the embedded language |specLangJS roundrobinf| is
by running multiple instances of the JS runtime, i.e., |targetLangJS|, in
separate OS-level threads.\footnote{
 The Firefox implementation of SWAPI also considers the case where
 tasks share the JavaScript runtime (and are cooperatively scheduled
 on the event-loop), but the heap and execution context of each task
 is disjoint. This is similar to the single-heap language of
 Section~\ref{sec:concrete}.  However, since the heap separation is
 provided by the browser the browser, we do not discuss this further.
 The interested reader is referred to~\tocite{} for more details.
}
%
In SWAPI, the IFC language functionality is implemented in the JS
runtime much like browser layout engines implement the DOM, and expose
the new functions, e.g., \verb|getElementById|, by attaching them to
the global JS object.
%
In turn, each task created with |sandbox| is
executed in a separate thread, running a separate, \emph{fresh} instance of our
modified JS runtime.
%
Formally, |sandbox| is defined with |klone(tS) =
tS0|, where |tS0| is the global object corresponding to the standard
JS library (e.g., |tS0| contains \texttt{Object}, \texttt{Array},
etc.).

Since this implementation approach relies on multiple runtimes of the
language, sending arbitrary objects between tasks can result in
unexpected behavior, e.g., when the object contains references.
%
As discussed in Section~\ref{sec:concrete}, we need to restrict |send|
to expressions that can be marshalled as strings, i.e., structurally
clonable objects.
%
In our formalization, this amounts to restricting the IFC language rule
for |send| such that only strings can be shared:
%{
\newcommand{\str}{"string"}
%format tOf (e) = "\texttt{typeOf}("e")\texttt{ === \str}"
%format ttrue = "\texttt{true}"
\begin{mathpar}
\inferrule[JS-send]
{
|il canFlowTo il'|\\
|iS(id') = Q|\\
|iS' = iS [ mapsto id' (il', id,  ie') , Q ]|\\
|ie = IT te|\\
|conf tS (tOf te) -> conf tS ttrue|
}
{|
iconf iS (fullconf id il tS (iniEi (send id' il' ie)), ldots)
.->
iS'; sched S (fullconf id il tS (iniEi unit), ldots)
|}
\end{mathpar}
%}
We remark that this is similar to the existing
\texttt{postMessage} API used for iframe and Worker
communications~\cite{webworkers}.
%
Thus, the SWAPI implementation provides an IFC version of
\texttt{postMessage}, defined in terms of |send|, that additionally
takes a label argument.
